---
title: "2.0 EDA"
author: "Andrew Ingalls"
date: "3/29/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse) #a little bit of everything all of the time
library(ggridges) #ridge plots
library(viridis) #coloring for ridge plots
library(ggResidpanel) # panel of diagnostic plots including cooksd
library(qqplotr) #qq plots
#library(EnvStats)
#library(lsmeans)
#library(plotly)
#library(outliers)
#library(hrbrthemes)
library(rstatix) #t-test, wilcoxon, anova, kruskal-wallis, correlation
#library(janitor)
#library(broom) #output tibbles for statistical modeling 
#library(datarium) #ANCOVA
library(ggpubr) #publication ready plots
library(ragg) #help with png creation
library(gridExtra) #allows for grid.arrange functionality

```

# Growth EDA
One of the three indicators for a successful feed is an improvement in growth. I 
use length and mass as the indicators. These are standard in the fish biology 
community. 

## Distribution Plots
First step in EDA is looking at distribution plots. What kind of distribution are
we working with? How do the age categories change the distribution? Do the 
distributions look different for length and mass? 

```{r Growth Distribution}



```

## Box plots 
```{r Box Plots}

```



# Survival EDA
:

```{r pressure, echo=FALSE}
plot(pressure)
```

# Sex Sort EDA
```{r}

```


# Fecundity EDA
```{r}

```

