---
title: "2.0 EDA"
author: "Andrew Ingalls"
date: "3/29/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)
library(viridis) #coloring for ridge plots
library(ragg) #help with png creation

```

```{r Constants for Graphing}

#Theme specifics for publication plots

theme_specs <- theme_minimal() +
  theme(
    axis.title = element_text(size = 30),
    axis.text = element_text(size = 20),
    plot.title = element_text(size = 40, face = "bold"),
    plot.subtitle = element_text(size = 25),
    legend.title = element_text(size =20),
    legend.text = element_text(size =15),
    legend.key.size = unit(2,"cm"),
    strip.text.x = element_text(size=20)
  ) 



```

# Growth EDA
One of the three indicators for a successful feed is an improvement in growth. I 
use length and mass as the indicators. These are standard in the fish biology 
community. We will explore distribution, box plots, scatter plots, outliers, etc.

## Distribution Plots
First step in EDA is looking at distribution plots. What kind of distribution are
we working with? How do the age categories change the distribution? Do the 
distributions look different for length and mass? 

```{r Growth Histograms}

#histogram for all mass data, including median line
mass_plot <- ggplot(df_growth, aes(x=Mass_mg)) 

mass_plot + 
  geom_histogram(aes(y=..density..)) +
  geom_vline(aes(xintercept = median(Mass_mg)), color="blue", 
             linetype="dashed", size=1) +
  geom_density(alpha=0.2, fill="#FF6666") +
  theme_minimal()

#histogram for all length data, including median line
length_plot <- ggplot(df_growth, aes(x=Length_mm)) 

length_plot + 
  geom_histogram(aes(y=..density..)) +
  geom_vline(aes(xintercept = median(Length_mm)), color="blue", 
             linetype="dashed", size=1) +
  geom_density(alpha=0.2, fill="#FF6666") +
  theme_minimal()
```

```{r Density Distributions by Groups}

density_specs <- geom_density(alpha=0.5, aes(colour =  Feed_Group, 
                                            fill = Feed_Group))

#Mass Density plots arranged by Feed Groups
Plot_Mass_Density_Feed_Groups <- mass_plot +
  density_specs +
  scale_color_brewer(palette="Dark2") + 
  scale_fill_brewer(palette="Dark2") + 
  labs(title='Mass Distribution of A. mexicanus by Feed Type', 
       subtitle = "Density Distribution",
       y = "Density", x = "Mass (mg)") +
  theme_specs

#Length Density plots arranged by Feed Groups
Plot_Length_Density_Feed_Groups <- length_plot +
  density_specs +
  scale_color_brewer(palette="Dark2") +
  scale_fill_brewer(palette="Dark2") +
  labs(title='Length Distribution of A. mexicanus by Feed Type', 
       subtitle = "Density Distribution", 
       y = "Density", x = "Length (mm)") +
  theme_specs

#Density Plot of Mass for Ages and Groups
Plot_Mass_Density_AgeandFeed <- mass_plot +
  density_specs + 
  facet_wrap(~Age_Category) +
  scale_color_brewer(palette="Dark2") +
  scale_fill_brewer(palette="Dark2") +
  labs(title='Mass Distribution of A. mexicanus by Age and Feed Type', 
       subtitle = "Density Distributions",
       y = "Density", x = "Mass (mg)") +
  theme_specs

#Density Plot of Length for Ages and Groups
Plot_Length_Density_AgeandFeed <- length_plot +
  density_specs + 
  facet_wrap(~Age_Category) +
  scale_color_brewer(palette="Dark2") + 
  scale_fill_brewer(palette="Dark2") + 
  labs(title='Density Distributions', 
       subtitle = "Length Distribution of A. mexicanus by Age and Feed Type",
       y = "Density", x = "Length (mm)") + 
  theme_specs


#Ridgeline distributions of length for ages and feed groups

ggplot(df_growth, aes(x=Length_mm,y=Age_Category, fill=..x..))+
  geom_density_ridges_gradient(scale=3, rel_min_height=0.01)+
  scale_fill_viridis(name="Length_mm",option="C")+
  labs(title='Length Distributions  of A. mexicanus by Age')+
  theme_minimal()+
  theme(
    legend.position = 'none',
    panel.spacing = unit(0.1,'lines'),
    strip.text.x = element_text(size=12)
  ) +
  facet_wrap(~Feed_Group,ncol=1)


#Ridgeline distributions of Mass for ages and feed groups
RidgeLine_Mass_AgeandFeed<- ggplot(df_growth, aes(x=Mass_mg, 
                                                  y=Age_Category, 
                                                  fill=Feed_Group)) +
  geom_density_ridges(scale=2, rel_min_height=0.01) + 
  labs(title='Mass Distributions of A. mexicanus by Age', 
       subtitle = "Mass of Surface Morph by Age", 
       y = "Age Category", x = "Mass (mg)") +
  theme_specs +
  facet_wrap(~Feed_Group,ncol=1) + 
  scale_color_brewer(palette="Dark2")

#Ridgeline Graph with 95% and 5% tails highlighted
ggplot(df_growth, aes(x=Mass_mg, y=Age_Category, 
                      fill=factor(stat(quantile)))) +
  geom_density_ridges_gradient(
    scale=2, 
    rel_min_height=0.01,
    calc_ecdf = TRUE,
    quantiles = c(0.025,0.975)
    ) +
  scale_fill_manual(
    name = "Probability", 
    values = c("#FF0000A0", "#A0A0A0A0", "#0000FFA0"),
    labels = c("(0, 0.025]", "(0.025, 0.975]", "(0.975, 1]")
    ) + 
  labs(title='Mass Distribution of A. mexicanus by Age', 
       subtitle = "5% and 95% density tails highlighted", 
       y = "Age Category", x = "Mass (mg)")+
  theme_minimal() +
  facet_wrap(~Feed_Group,ncol=1) + 
  scale_color_brewer(palette="Dark2")
```

#Normal Distribution QQ Plot
Since some of of our analyses may require us to assume normality in the future, 
let's take a look at some normal Q-Q plots to see if any of the data does line up 
in a normal distribution. To do this we will iterate through the age levels of the database
and create a normal Q-Q plot for each age group using qqplotr and ggplot

```{r Normal Q-Q Plot for each Feed Group per Age Category for Mass and Length}

####Mass####
for (FeedGroup in levels(df_growth$Feed_Group)) {
  FeedGroup_df <-df_growth %>%
  filter(Feed_Group == FeedGroup)
  for (AgeCategory in levels(FeedGroup_df$Age_Category)){
    AgeGroup_df <- FeedGroup_df %>%
    filter(Age_Category == AgeCategory)
   
    QQPlot_AgeGroup <- ggplot(mapping = aes(sample =AgeGroup_df$Mass_mg)) + 
      stat_qq_point(size = 2) + 
      stat_qq_line(color = "green") + 
      xlab("Normal N(0,1) Order Statistic Medians") +
      ylab("Ordered Response") + 
      labs(title = paste(FeedGroup,"Mass Normal Probabilty Plot:", AgeCategory))
    print(QQPlot_AgeGroup) #Will print out a plot for each age group iteration
  }
}

####Length####
for (FeedGroup in levels(df_growth$Feed_Group)) {
  FeedGroup_df <-df_growth %>%
  filter(Feed_Group == FeedGroup)
  for (AgeCategory in levels(FeedGroup_df$Age_Category)){
    AgeGroup_df <- FeedGroup_df %>%
    filter(Age_Category == AgeCategory)
  
    QQPlot_AgeGroup <- ggplot(mapping = aes(sample =AgeGroup_df$Length_mm)) + 
      stat_qq_point(size = 2) +
      stat_qq_line(color = "green") + 
      xlab("Normal N(0,1) Order Statistic Medians") +
      ylab("Ordered Response") + 
      labs(title = paste(FeedGroup,"Length Normal Probabilty Plot:", AgeCategory))
    print(QQPlot_AgeGroup)
  }
}
```

## Growth Box plots
Now that we have determined the distributions of out growth dataset, we will turn 
to box plots to compare both length and mass between groups, ages, and tank locations

```{r Growth Box Plots}
#constants in box plots
mass_age_plots <- ggplot(df_growth,aes(y=Mass_mg, x=Age_Category))
length_age_plots <- ggplot(df_growth,aes(y=Length_mm, x=Age_Category))
boxplot_specs <- geom_boxplot(aes(color=Feed_Group), 
                             lwd=2, 
                             fill='grey', 
                             outlier.colour = NULL, 
                             outlier.size = 3) 

#Box plot of Mass vs Age for Groups

Mass_boxplots_AgeandFeed <-  mass_age_plots + 
  boxplot_specs +
  scale_color_brewer(palette="Dark2") +
  labs(title="Mass of A. mexicanus by Age & Feed Type", 
       subtitle = 'Box Plot' ,
       y = "Mass (mg)", x = "Age Category") +
  theme_specs

#Box plot of Length vs Age for Groups

Length_boxplots_AgeandFeed <- length_age_plots +
  boxplot_specs +
  scale_color_brewer(palette="Dark2") +
  labs(title='Length of A. mexicanus by Age & Feed Type', 
       subtitle = "Box Plot",
       y = "Length (mm)", x = "Age Category") +
  theme_specs


#Violin plots
mass_age_plots + 
  geom_violin(aes(fill=Feed_Group)) + 
  stat_summary(fun=median, geom='point', size=2, color='red') +
  facet_wrap(~Feed_Group) + 
  theme_minimal() + 
  theme(axis.text.x=element_text(angle=45,hjust=1))

length_age_plots + 
  geom_violin(aes(fill=Feed_Group)) + 
  stat_summary(fun=median,geom='point',size=2,color='red') +
  facet_wrap(~Feed_Group) +
  theme_minimal() + 
  theme(axis.text.x=element_text(angle=45,hjust=1))

#Understanding value spread based on Age category. 
mass_age_plots + 
  geom_point(alpha=0.5,aes(color=Feed_Group))
length_age_plots + 
  geom_point(alpha=0.5,aes(color=Feed_Group))


#understanding Mass of each Feed Group through tank locations and Age
  mass_age_plots +
  geom_boxplot(aes(color=Age_Category)) +
  theme_minimal() +
  theme(legend.position = 'none',
        axis.text.x = element_text(angle=45,
                                   hjust=1)) +
  facet_wrap(Tank_Location~Feed_Group,ncol=4)

#looking at growth variables vs age 
#through the lens of our different tank locations and groups

Mass_tanks_age_boxplots <- mass_age_plots + 
  geom_boxplot(lwd=1,
               fill = 'grey',
               aes(color= Tank_Location), 
               outlier.colour = NULL,
               outlier.size = 2) +
  scale_color_brewer(palette="Dark2") +
  labs(title='Mass by Age, Tank Location, & Feed Type', 
       subtitle = "Box Plot",
       y = "Mass (mg)", x = "Age Category") +
  theme_specs +
  facet_wrap(~Feed_Group,ncol=1)

Length_tanks_age_boxplots <- length_age_plots + 
  geom_boxplot(lwd=1,
               fill = 'grey',
               aes(color= Tank_Location), 
               outlier.colour = NULL,
               outlier.size = 2) +
  scale_color_brewer(palette="Dark2") +
  labs(title='Length by Age, Tank Location, & Feed Type', 
       subtitle = "Box Plot",
       y = "Length (mm)", x = "Age Category") +
  theme_specs +
  facet_wrap(~Feed_Group,ncol=1)


```

## Growth: Scatter Analysis
Now that we have an understanding of the distribution, spread, and differences between
the feed groups for length and mass, it's time to look at how the two values correlate
to each other. We will look at Feed group and age to see if the relationship has
any visible grouping

```{r Growth Length vs Mass Relationship Scatter Plots}

#Feed Groups as colors
Length_Mass_Relationship_Group <- ggplot(df_growth, 
                                         aes(x=Length_mm, 
                                             y=Mass_mg, 
                                             color=Feed_Group)) + 
  geom_point(size=2, alpha=0.7) +
  scale_color_brewer(palette="Dark2") +
  labs(title='Mass vs Length Relationship', 
       subtitle = 'Feed Groups as colors',
       y = "Mass (mg)", x = "Length (mm)") +
  theme_specs

#Age as Colors and Feed Group as Shape
Length_Mass_Relationship_Age <- ggplot(df_growth, 
                                       aes(x=Length_mm, 
                                           y=Mass_mg, 
                                           color=Age_Category)) + 
  geom_point(size=2, alpha=0.7, 
             aes(shape=Feed_Group)) +
  scale_color_brewer(palette="YlGnBu", direction = -1) +
  labs(title='Mass vs Length Relationship Feed',
       subtitle='Age Categories',
       y = "Mass (mg)", x = "Length (mm)") +
  theme_dark() +
  theme(
    axis.title = element_text(size = 30),
    axis.text = element_text(size = 20),
    plot.title = element_text(size = 40, face = "bold"),
    plot.subtitle = element_text(size = 25),
    legend.title = element_text(size =25),
    legend.text = element_text(size =20),
    legend.key.size = unit(2,"cm"),
    strip.text.x = element_text(size=20)
    )

```


# Survival EDA
The second of the three indicators for a successful feed, Surival is of particular
importance for A. mexicanus. The surface morph (studied here) has a high rate
of aggression related attacks and deaths. Studying the impact feeds have on the surivival
during the high aggression periods (three to six months old) can help minimize death
and maximize the welfare of the animals. 

##Time Series Analysis
Since census was taken at certain points over a period of nine months, time series
line plots will be the best way for us to understand the trends in surival over time

```{r Survival Line plots}
#individual tank location census over time grouped by feed type
Time_series_location_feed <- ggplot(df_survival, 
                                    aes(x=Date, y=Census, 
                                        group = Tank_Location, 
                                        color = Feed_Group)) +
  geom_line(size=2) + 
  geom_point(size=4) +
  scale_color_brewer(palette="Dark2") +
  labs(title='Fish Census per Tank Location over Time', 
       y = "Fish Census", x = "Date") +
  theme_specs

#Summarize the data by feedgroup and date, using average
#Summarize standard error of census per group per timepoint
#plot census with standard error against time using feed groups
Summary_time_series_feed <- df_survival %>%
  group_by(Feed_Group,Date) %>%
  summarise(Census_avg=mean(Census), 
            Standard_Error=sd(Census)/sqrt(length(Census))) %>%
  ggplot(.,aes(x=Date,y=Census_avg,color=Feed_Group)) +geom_line(size=2)+
  geom_point(size = 4)+ geom_errorbar(aes(ymin=Census_avg-Standard_Error,
                                  ymax=Census_avg+Standard_Error),width=5) +
  scale_color_brewer(palette="Dark2") +
  labs(title='Average Feed Group Census over Time', 
       subtitle = 'Standard Error for Comparison',
       y = "Fish Census", x = "Date") +
  theme_specs

```

## Barplots for Average Death
While the line plots show us how the census changed over time, I am also interested 
in understanding the number of deaths from event to event, and how that changes.
It is possible that there is one large death event in the beginning that separates 
the two groups, then each census following has the same slope. Showing the average deaths
between events will tell us if the two groups consistently differ.

```{r Survival Bar Plot Deaths over Time}
#plot the difference in means between groups from one census event to the next
#e.g. first bars are 1st Census - 2nd Census (20-14)
Barplot_avgdeaths_feed<- df_survival %>%
  group_by(Date, Feed_Group) %>%
  summarize(Census = mean(Census)) %>%
  group_by(Feed_Group) %>%
  summarize(Census_Dead = -diff(Census),
            Date = Date[-1]) %>%
   ggplot(aes(y = Census_Dead, x= Date)) +
  geom_col(aes(fill = Feed_Group), position = "dodge") +
  scale_color_brewer(palette="Dark2") +
  labs(title='Average Feed Group Deaths between Census Events', 
       y = "Fish Deaths", x = "Date") +
  theme_specs
```

# Sex Sort EDA
While not a primary indicator for the success of a feed, Sex skews are very common
in animal husbandry. They can be caused by environment, diets, health, or a number
of other stressors. This has a large role in both the facility (how many fish do
we need to rear in order to get X number of males and females) as well as indirectly
impacting the fecundity (Will there be enough males and females for the proper ratio)

```{r Sex Sort EDA}
#Create a column of total counts for each tank location
df_sex_sort$Total <- rowSums(df_sex_sort[,c("Female","Male","Unknown")])
#normalize each tank's sex data between 1 and 0 using total counts
df_sex_norm <- df_sex_sort %>%
  group_by(Feed_Group)%>%
  summarise(Female=Female/Total,Male=Male/Total,Unknown=Unknown/Total)

#tidy data (long), summarize for mean and standard error
df_sex_mean<-df_sex_norm %>%
   pivot_longer(.,Female:Unknown, names_to = "sex",values_to = "Census") %>%
  group_by(Feed_Group, sex) %>%
  summarise(Mean=mean(Census),Standard_Error=sd(Census)/sqrt(length(Census)))

#create bar plot of sex by group with standard error  
 Sex_sort_Barplot <- ggplot(df_sex_mean,aes(x=sex,y=Mean,fill=Feed_Group)) + 
   geom_col(position="dodge2") +
   geom_errorbar(aes(ymin=Mean-Standard_Error,
                     ymax=Mean+Standard_Error),position="dodge2") +
  scale_color_brewer(palette="Dark2") +
  labs(title='Fish Sex Sort by Groups', 
       y = "Percent of Group", x = "Sex") +
  theme_specs
  
```


# Fecundity EDA
The final indicator of a successful feed, fecundity is important to the long term
success of our colony as well as embryo generation reliability for our research. 
I will look at the total viable embryos for each tank. Viable embryos are those still
alive at the time of collection (~14 hours post fertilization). This can be different
than total embryos, as some are not fertilized or die off before they start to
grow. I will also be looking at the average viable production for each group, using
standard error as a visual for significance. 

```{r Fecundity EDA}
#Chart of each tank location and the total of all viable embryos produced
Sum_Viable_Embryos_Tank <- ggplot(df_fecundity, 
                                  aes(x=Sample_ID, 
                                      Viable_Embryos, 
                                      fill=Feed_Group)) + 
  geom_col() +
  scale_color_brewer(palette="Dark2") +
  labs(title='Total Viable Embryo Production', 
       subtitle = 'Over a Three Month Period',
       y = "Total Viable Embryo Count", x = "Sample Number") +
  theme_specs

#After filtering out 10C6 (extra tank), the average viable production and standard 
#error for each group
Average_Viable_Plot<-df_fecundity %>%
  filter(Sample_ID != "10C6")%>% #remove 10C6
  group_by(Feed_Group) %>%
  summarize(Mean = mean(Viable_Embryos),
            Standard_Error =
              sd(Viable_Embryos)/sqrt(length(Viable_Embryos))) %>%
  ggplot(., aes(x=Feed_Group, y=Mean, fill=Feed_Group)) + 
  geom_col() + 
  geom_errorbar(aes(ymin = Mean-Standard_Error,
                    ymax = Mean+Standard_Error)) +
  scale_color_brewer(palette="Dark2") +
  labs(title='Average Viable Embryo Production', 
       subtitle = 'Over a Three Month Period',
       y = "Average Viable Embryo Count", x = "Feed Group") +
  theme_specs

```



# Printing pngs for Medium posts

```{r Medium plot saves}

####Growth EDA Plots####
pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/density_mass.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Plot_Mass_Density_Feed_Groups)
invisible(dev.off())
knitr::include_graphics(pngfile)

pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/density_length.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Plot_Length_Density_Feed_Groups)
invisible(dev.off())
knitr::include_graphics(pngfile)

pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/Density_Age_Mass.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Plot_Mass_Density_AgeandFeed)
invisible(dev.off())
knitr::include_graphics(pngfile)

pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/Density_Age_Length.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Plot_Length_Density_AgeandFeed)
invisible(dev.off())
knitr::include_graphics(pngfile)

pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/BoxPlot_Mass.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Mass_boxplots_AgeandFeed)
invisible(dev.off())
knitr::include_graphics(pngfile)

pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/BoxPlot_Length.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Length_boxplots_AgeandFeed)
invisible(dev.off())
knitr::include_graphics(pngfile)

pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/BoxPlot_Tanks_Mass.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Mass_tanks_age_boxplots)
invisible(dev.off())
knitr::include_graphics(pngfile)

pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/BoxPlot_Tanks_Length.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Length_tanks_age_boxplots)
invisible(dev.off())
knitr::include_graphics(pngfile)

pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/LMR_SP.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Length_Mass_Relationship_Group)
invisible(dev.off())
knitr::include_graphics(pngfile)

pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/LMR_SP_Age.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Length_Mass_Relationship_Age)
invisible(dev.off())
knitr::include_graphics(pngfile)


####Survival Plots####
pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/Survival_Tanks.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Time_series_location_feed)
invisible(dev.off())
knitr::include_graphics(pngfile)

pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/Survival_Avg.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Summary_time_series_feed)
invisible(dev.off())
knitr::include_graphics(pngfile)

pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/Deaths_Bar.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Barplot_avgdeaths_feed)
invisible(dev.off())
knitr::include_graphics(pngfile)


####Sex Sort Plots####
pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/Sex_sort_mean.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Sex_sort_Barplot)
invisible(dev.off())
knitr::include_graphics(pngfile)

####Fecundity Plots####

pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/Total_Viable.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Sum_Viable_Embryos_Tank)
invisible(dev.off())
knitr::include_graphics(pngfile)

pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/Average_Viable.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Average_Viable_Plot)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

