---
title: "3.0 Outliers"
author: "Andrew Ingalls"
date: "3/29/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggResidpanel) # panel of diagnostic plots including cooksd
library(qqplotr)
library(EnvStats) #RosnerTest
library(ragg) #help with png creation

```

# Create Outlier Dataset
To preserve the dataframes we are working with, I will create a series of copy sets
in order to look at outliers. Split the data into Gemma and Mysis.

```{r Outlier Dataset}

growth_gemma_outlier_df <- df_growth %>% filter(Feed_Group == "Gemma")

growth_mysis_outlier_df <- df_growth %>% filter(Feed_Group == "Mysis")

summary(growth_mysis_outlier_df)
summary(growth_gemma_outlier_df)

```


# Mass Outliers for Mysis and Gemma
 I will look at Mass  outliers for both Mysis and Gemma feed groups. There are 
 four methods of outlier detection I will use to look at outliers: Boxplots, 
 percentile, Rosner outliers, and Cook's distance. I am doing this more for
 curiosity sake as we will keep all the outliers, since they are expected. Additionally,
 Rosners is not a great detection method here since we don't have normality with 
 all the data. 

## Boxplot Outliers

```{r Boxplot Outliers }

#Boxplot outlier statistics
mysis_mass_outliers <- growth_mysis_outlier_df %>%
  group_by(Age_Category) %>%
  summarize(outliers = boxplot.stats(Mass_mg)$out,.groups = 'drop') %>%
  unnest_wider(outliers)

gemma_mass_outliers <- growth_gemma_outlier_df %>%
  group_by(Age_Category) %>% 
  summarize(outliers = boxplot.stats(Mass_mg)$out,.groups = 'drop') %>%
  unnest_wider(outliers)

#Create dataframe of boxplot outliers for each Feed Group
mysis_mass_boxplot_outlier_df <- data.frame()
gemma_mass_boxplot_outlier_df <- data.frame()

#match Age Category and Mass Value from boxplot outlier df  
#with a row in the original outlier dataframe and copy it to new dataframe.
for (i in 1:nrow(mysis_mass_outliers)) {
 OneRowofOutlier <- growth_mysis_outlier_df %>%
 filter(.,Age_Category == mysis_mass_outliers[[i,"Age_Category"]]) %>%
   filter(., Mass_mg == mysis_mass_outliers[[i,"...1"]])
 
 mysis_mass_boxplot_outlier_df <- rbind(mysis_mass_boxplot_outlier_df, 
                                        OneRowofOutlier)
}

for (i in 1:nrow(gemma_mass_outliers)) {
 OneRowofOutlier <- growth_gemma_outlier_df %>%
 filter(.,Age_Category == gemma_mass_outliers[[i,"Age_Category"]]) %>%
   filter(., Mass_mg == gemma_mass_outliers[[i,"...1"]])
 
 gemma_mass_boxplot_outlier_df <- rbind(gemma_mass_boxplot_outlier_df, 
                                        OneRowofOutlier)
}


mysis_mass_boxplot_outlier_df <- mysis_mass_boxplot_outlier_df %>% distinct()
gemma_mass_boxplot_outlier_df <- gemma_mass_boxplot_outlier_df %>% distinct()


```

## Percentile Outliers
Percentile outliers will flag any sample outside 2.5 and 97.5 percentiles. First 
I will need to find the upper and lower percentile bounds. 

```{r Percentile Outliers }
mysis_lower_bound <- growth_gemma_outlier_df %>%
  group_by(Age_Category) %>%
  summarize(lower_quantile = quantile(Mass_mg, 0.025))

mysis_upper_bound <- growth_gemma_outlier_df %>%
  group_by(Age_Category) %>%
  summarize(upper_quantile = quantile(Mass_mg, 0.975))

gemma_lower_bound <- growth_gemma_outlier_df %>%
  group_by(Age_Category) %>%
  summarize(lower_quantile = quantile(Mass_mg, 0.025))

gemma_upper_bound <- growth_gemma_outlier_df %>%
  group_by(Age_Category) %>%
  summarize(upper_quantile = quantile(Mass_mg, 0.975))


#percentile df of the outliers
mysis_mass_percentile_outlier_df <- data.frame()
gemma_mass_percentile_outlier_df <- data.frame()

#iterate through all age categories filtering any values outside the category's bounds
for (AgeCategory in levels(growth_mysis_outlier_df$Age_Category)){
    OneRowofOutlier <- growth_mysis_outlier_df %>%
    filter(Age_Category == AgeCategory) %>%
    filter(Mass_mg < 
             mysis_lower_bound$lower_quantile[mysis_lower_bound$Age_Category == AgeCategory] |
           Mass_mg > 
             mysis_upper_bound$upper_quantile[mysis_lower_bound$Age_Category == AgeCategory])
    
    mysis_mass_percentile_outlier_df  <-rbind(mysis_mass_percentile_outlier_df, OneRowofOutlier)
}

for (AgeCategory in levels(growth_gemma_outlier_df$Age_Category)){
  OneRowofOutlier <- growth_gemma_outlier_df %>%
    filter(Age_Category == AgeCategory) %>%
    filter(Mass_mg < 
             gemma_lower_bound$lower_quantile[gemma_lower_bound$Age_Category==AgeCategory] | 
           Mass_mg > 
             gemma_upper_bound$upper_quantile[gemma_upper_bound$Age_Category==AgeCategory])
  
  gemma_mass_percentile_outlier_df <-rbind(gemma_mass_percentile_outlier_df, OneRowofOutlier)
}

```

## Generalized ESD Test for Outliers
Generalized ESD test are used to detect one or more outliers in a univariate dataset
for approximate normal distribution (removing the influence of the outliers). 
We do not have to specify the number of outliers, just an upper bound. 

```{r Rosner GESD Test for Mass Outliers }
mysis_mass_rosner_outlier_df <- data.frame()
gemma_mass_rosner_outlier_df <- data.frame()

#iterate through age categories and create temp df for that age category
for (AgeCategory in levels(growth_mysis_outlier_df$Age_Category)){
  AgeGroup_df <- growth_mysis_outlier_df %>%
  filter(Age_Category == AgeCategory)
  mysis_rosner_mass <- rosnerTest(AgeGroup_df$Mass_mg, k=10, alpha=0.05) #10 outliers max

  Rosner_Outliers <- mysis_rosner_mass[["all.stats"]] %>%
    filter(Outlier == TRUE) 

  mysis_rosner_outlier_rownum <- AgeGroup_df %>%  
                                  filter(row_number() %in% Rosner_Outliers$Obs.Num) #Obs.Num = row

  mysis_mass_rosner_outlier_df <- rbind(mysis_mass_rosner_outlier_df, 
                                        mysis_rosner_outlier_rownum)
}

for (AgeCategory in levels(growth_gemma_outlier_df$Age_Category)){
  AgeGroup_df <- growth_gemma_outlier_df %>%
  filter(Age_Category == AgeCategory)
  gemma_rosner_mass <- rosnerTest(AgeGroup_df$Mass_mg, k=10, alpha=0.05)

  Rosner_Outliers <- gemma_rosner_mass[["all.stats"]] %>%
    filter(Outlier == TRUE)

  gemma_rosner_outlier_rownum <- AgeGroup_df %>% 
                                    filter(row_number() %in% Rosner_Outliers$Obs.Num)

  gemma_mass_rosner_outlier_df <- rbind(gemma_mass_rosner_outlier_df, 
                                        gemma_rosner_outlier_rownum)
}

```


## Cooks Distance Outliers
Estimates the amount of influence each observation has on the over all 
regression model. It does this by removing each observation and 
calculating the change in the model. The plots aren't necessary, but I find them 
helpful when looking at cooks distance

```{r Cooks Distance Outliers}
mysis_mass_cooksd_outlier_df <- data.frame()
gemma_mass_cooksd_outlier_df <- data.frame()


for (AgeCategory in levels(growth_gemma_outlier_df$Age_Category)){
  AgeGroup_df <- growth_gemma_outlier_df %>%
  filter(Age_Category == AgeCategory)

  model <- lm(Mass_mg ~ Length_mm, data = AgeGroup_df)
  cooksD <- cooks.distance(model)
  influence_cuttoff <- 3 * mean(cooksD, na.rm =TRUE)
  influencial_Observations <- cooksD[(cooksD > influence_cuttoff)]
 
  plot(cooksD, pch="*", cex=2, main = "Influential Obs by Cooks distance") 
  abline(h=influence_cuttoff, col="red") 
  text(x = 1:length(cooksD) + 1, 
       y = cooksD, 
       labels = ifelse(cooksD > influence_cuttoff, names(cooksD),""), 
       col = "red") 
 
  gemma_cooksd_outlier_rownum <- AgeGroup_df %>% 
                                filter(row_number() %in% 
                                         names(influencial_Observations))

  gemma_mass_cooksd_outlier_df <- rbind(gemma_mass_cooksd_outlier_df, 
                                       gemma_cooksd_outlier_rownum)
 
}


for (AgeCategory in levels(growth_mysis_outlier_df$Age_Category)){
  AgeGroup_df <- growth_mysis_outlier_df %>%
  filter(Age_Category == AgeCategory)

  model <- lm(Mass_mg ~ Length_mm, data = AgeGroup_df)
  cooksD <- cooks.distance(model)
  influencial_observations <- cooksD[(cooksD > influence_cuttoff)]
 
 plot(cooksD, pch="*", cex=2, main = "Influential Obs by Cooks distance") 
 abline(h = influence_cuttoff,col="red")
 text(x=1:length(cooksD) + 1, 
      y=cooksD, 
      labels = ifelse(cooksD > influence_cuttoff, names(cooksD),""),
      col="red")
 
  mysis_cooksd_outlier_rownum <- AgeGroup_df %>% 
                                filter(row_number() %in% 
                                         names(influencial_observations))

  mysis_mass_cooksd_outlier_df <- rbind(mysis_mass_cooksd_outlier_df, 
                                        mysis_cooksd_outlier_rownum)
 
 
}

```

## Mass Outlier Comparison
This section will be comparing all the different methods of outlier detection 
and the overlap in outliers detected. What are the similarities and differences?


```{r Mass Outlier Summary}

#create summary dataframe of all possible outliers using a list of dataframes for antijoin
mysis_mass_summary_outlier_df <- mysis_mass_boxplot_outlier_df
gemma_mass_summary_outlier_df <- gemma_mass_boxplot_outlier_df

mysis_list_outlier_df <- list(mysis_mass_boxplot_outlier_df, mysis_mass_percentile_outlier_df,
                              mysis_mass_rosner_outlier_df, mysis_mass_cooksd_outlier_df)
gemma_list_outlier_df <- list(gemma_mass_boxplot_outlier_df, gemma_mass_percentile_outlier_df,
                              gemma_mass_rosner_outlier_df, gemma_mass_cooksd_outlier_df)


for (outlier_df in mysis_list_outlier_df) {
  Outliers_notin_summary_df <- anti_join(outlier_df, mysis_mass_summary_outlier_df)

  mysis_mass_summary_outlier_df <- rbind(mysis_mass_summary_outlier_df, 
                                        Outliers_notin_summary_df)
}

for (outlier_df in gemma_list_outlier_df) {
  Outliers_notin_summary_df <- anti_join(outlier_df, gemma_mass_summary_outlier_df)

  gemma_mass_summary_outlier_df <- rbind(gemma_mass_summary_outlier_df, 
                                        Outliers_notin_summary_df)
}

#Wrangle Dataframes
mysis_mass_summary_outlier_df <- mysis_mass_summary_outlier_df %>% arrange(Sample_ID)
gemma_mass_summary_outlier_df <- gemma_mass_summary_outlier_df %>% arrange(Sample_ID)

row.names(mysis_mass_summary_outlier_df) <- NULL
row.names(gemma_mass_summary_outlier_df) <- NULL

mysis_mass_outlier_method_df <- subset(mysis_mass_summary_outlier_df,
                                        select=-c(Date, Tank_Location, Feed_Group, Length_mm)) 
gemma_mass_outlier_method_df <- subset(gemma_mass_summary_outlier_df,
                                        select=-c(Date, Tank_Location, Feed_Group, Length_mm)) 

#add columns for each method
methods_list = list("Boxplot", "Percentile", "Rosner", "CooksD")

for (method in methods_list){
  mysis_mass_outlier_method_df <- mysis_mass_outlier_method_df %>% 
    add_column(!!(method) :=0) #add column for each list and set value to 0
  gemma_mass_outlier_method_df <- gemma_mass_outlier_method_df %>%
    add_column(!!(method) :=0) 
}


# for each outlier, add +1 to any method column that detected that outlier. 
###Mysis###
for (sample_ID in mysis_mass_outlier_method_df$Sample_ID){
  if (sample_ID %in% mysis_mass_boxplot_outlier_df$Sample_ID){
    mysis_mass_outlier_method_df["Boxplot"][mysis_mass_outlier_method_df["Sample_ID"] == 
                                              sample_ID] <- 1
  }
  if (sample_ID %in% mysis_mass_percentile_outlier_df$Sample_ID){
    mysis_mass_outlier_method_df["Percentile"][mysis_mass_outlier_method_df["Sample_ID"] ==
                                                 sample_ID] <-1
  }
  if (sample_ID %in% mysis_mass_rosner_outlier_df$Sample_ID){
    mysis_mass_outlier_method_df["Rosner"][mysis_mass_outlier_method_df["Sample_ID"] == 
                                             sample_ID] <-1
  }
  if (sample_ID %in% mysis_mass_cooksd_outlier_df$Sample_ID){
    mysis_mass_outlier_method_df["CooksD"][mysis_mass_outlier_method_df["Sample_ID"] == 
                                             sample_ID] <-1
  }
}

###Gemma###
for (sample_ID in gemma_mass_outlier_method_df$Sample_ID){
  if (sample_ID %in% gemma_mass_boxplot_outlier_df$Sample_ID){
    gemma_mass_outlier_method_df["Boxplot"][gemma_mass_outlier_method_df["Sample_ID"] == 
                                              sample_ID] <-1
  }
  if (sample_ID %in% gemma_mass_percentile_outlier_df$Sample_ID){
    gemma_mass_outlier_method_df["Percentile"][gemma_mass_outlier_method_df["Sample_ID"] ==
                                                 sample_ID] <-1
  }
  if (sample_ID %in% gemma_mass_rosner_outlier_df$Sample_ID){
    gemma_mass_outlier_method_df["Rosner"][gemma_mass_outlier_method_df["Sample_ID"] == 
                                             sample_ID] <-1
  }
  if (sample_ID %in% gemma_mass_cooksd_outlier_df$Sample_ID){
    gemma_mass_outlier_method_df["CooksD"][gemma_mass_outlier_method_df["Sample_ID"] == 
                                             sample_ID] <-1
  }
}

#Create a total column of outlier methods, to sum all detections among methods per sample

mysis_mass_outlier_method_df <- mysis_mass_outlier_method_df %>%
 mutate(Total = select(., Boxplot:CooksD) %>% rowSums(na.rm =TRUE))

gemma_mass_outlier_method_df <- gemma_mass_outlier_method_df %>%
 mutate(Total = select(., Boxplot:CooksD) %>% rowSums(na.rm =TRUE))

#Add a Mass_outlier column to Mysis Outlier Dataframe. Transfer Total Method data 
#to Mass outlier samples. Insert 0, where no outliers detected.
for (Sample in growth_mysis_outlier_df$Sample_ID) {
  if (Sample %in% mysis_mass_outlier_method_df$Sample_ID){
    growth_mysis_outlier_df$Mass_Outlier[growth_mysis_outlier_df["Sample_ID"] == Sample] <- 
      mysis_mass_outlier_method_df$Total[mysis_mass_outlier_method_df["Sample_ID"] == Sample]
  } 
  else{
    growth_mysis_outlier_df$Mass_Outlier[growth_mysis_outlier_df["Sample_ID"] == Sample] <- 0
  }
}

for (Sample in growth_gemma_outlier_df$Sample_ID) {
  if (Sample %in% gemma_mass_outlier_method_df$Sample_ID){
    growth_gemma_outlier_df$Mass_Outlier[growth_gemma_outlier_df["Sample_ID"] == Sample] <- 
      gemma_mass_outlier_method_df$Total[gemma_mass_outlier_method_df["Sample_ID"] == Sample]
  } 
  else{
    growth_gemma_outlier_df$Mass_Outlier[growth_gemma_outlier_df["Sample_ID"] == Sample] <- 0
  }
}

growth_mysis_outlier_df$Mass_Outlier <- factor(growth_mysis_outlier_df$Mass_Outlier, 
                                 levels = c("4","3","2","1","0"))

growth_gemma_outlier_df$Mass_Outlier <- factor(growth_gemma_outlier_df$Mass_Outlier, 
                                 levels = c("4","3","2","1","0"))



#plot outlier frequency based on method selection
Mysis_LMR_Scatter_Mass_Outliers <- ggplot(growth_mysis_outlier_df, 
       aes(x = Length_mm, y = Mass_mg, color = Mass_Outlier)) +
  geom_point()+
  scale_color_brewer(palette="Set1") 

Gemma_LMR_Scatter_Mass_Outliers <- ggplot(growth_gemma_outlier_df, 
       aes(x = Length_mm, y = Mass_mg, color = Mass_Outlier)) +
  geom_point()+
  scale_color_brewer(palette="Set1") 


Mysis_Mass_Boxplot_Mass_Outliers <- ggplot(growth_mysis_outlier_df, 
       aes(y=Mass_mg, x=Age_Category)) + 
  geom_boxplot() + 
  geom_point(aes(color = Mass_Outlier))+
  scale_color_brewer(palette="Set1") + theme_classic()

Gemma_Mass_Boxplot_Mass_Outliers <- ggplot(growth_gemma_outlier_df, 
       aes(y=Mass_mg, x=Age_Category)) + 
  geom_boxplot() + 
  geom_point(aes(color = Mass_Outlier))+
  scale_color_brewer(palette="Set1") + theme_classic()

```


#Length Outliers for Mysis and Gemma

## Boxplot Outliers

```{r Length Outlier Boxplot}

#Boxplot outlier statistics
mysis_length_outliers <- growth_mysis_outlier_df %>% 
  group_by(Age_Category) %>%
  summarize(outliers = boxplot.stats(Length_mm)$out, .groups = 'drop') %>%
  unnest_wider(outliers)

gemma_length_outliers <- growth_gemma_outlier_df %>%
  group_by(Age_Category) %>%
  summarize(outliers =boxplot.stats(Length_mm)$out) %>%
  unnest_wider(outliers)


#Create dataframe of boxplot outliers for each Feed Group
mysis_length_boxplot_outlier_df <- data.frame()
gemma_length_boxplot_outlier_df <- data.frame()

#match Age Category and Mass Value from boxplot outlier df  
#with a row in the original outlier dataframe and copy it to new dataframe.
for (i in 1:nrow(mysis_length_outliers)) {
 OneRowofOutlier <- growth_mysis_outlier_df %>%
 filter(., Age_Category == mysis_length_outliers[[i, "Age_Category"]]) %>%
   filter(., Length_mm == mysis_length_outliers[[i, "...1"]])
 
 mysis_length_boxplot_outlier_df <- rbind(mysis_length_boxplot_outlier_df,
                                          OneRowofOutlier)
}

for (i in 1:nrow(gemma_length_outliers)) {
  OneRowofOutlier <- growth_gemma_outlier_df %>%
    filter(., Age_Category == gemma_length_outliers[[i, "Age_Category"]]) %>%
    filter(., Length_mm == gemma_length_outliers[[i, "...1"]])
  
 gemma_length_boxplot_outlier_df <- rbind(gemma_length_boxplot_outlier_df,
                                          OneRowofOutlier)
}

mysis_length_boxplot_outlier_df <- mysis_length_boxplot_outlier_df %>% distinct()
gemma_length_boxplot_outlier_df <- gemma_length_boxplot_outlier_df %>% distinct()

```

## Percentile Outliers
```{r Length Outliers Percentiles}
mysis_lower_bound <- growth_gemma_outlier_df %>%
  group_by(Age_Category) %>%
  summarize(lower_quantile = quantile(Length_mm, 0.025))

mysis_upper_bound <- growth_gemma_outlier_df %>%
  group_by(Age_Category) %>%
  summarize(upper_quantile = quantile(Length_mm, 0.975))

gemma_lower_bound <- growth_gemma_outlier_df %>%
  group_by(Age_Category) %>%
  summarize(lower_quantile = quantile(Length_mm, 0.025))

gemma_upper_bound <- growth_gemma_outlier_df %>%
  group_by(Age_Category) %>%
  summarize(upper_quantile = quantile(Length_mm, 0.975))

#percentile df of the outliers
gemma_length_percentile_outlier_df <- data.frame()
mysis_length_percentile_outlier_df <- data.frame()

#iterate through all age categories filtering any values outside the category's bounds
for (AgeCategory in levels(growth_gemma_outlier_df$Age_Category)){
  OneRowofOutlier <- growth_gemma_outlier_df %>%
    filter(Age_Category == AgeCategory) %>%
    filter(Length_mm < 
             gemma_lower_bound$lower_quantile[gemma_lower_bound$Age_Category == AgeCategory] | 
           Length_mm > 
             gemma_upper_bound$upper_quantile[gemma_upper_bound$Age_Category == AgeCategory])
  gemma_length_percentile_outlier_df <-rbind(gemma_length_percentile_outlier_df,
                                             OneRowofOutlier)
  }
  

for (AgeCategory in levels(growth_mysis_outlier_df$Age_Category)){
  OneRowofOutlier <- growth_mysis_outlier_df %>%
  filter(Age_Category == AgeCategory) %>%
  filter(Length_mm < 
             mysis_lower_bound$lower_quantile[mysis_lower_bound$Age_Category == AgeCategory] | 
           Length_mm > 
             mysis_upper_bound$upper_quantile[mysis_upper_bound$Age_Category == AgeCategory])
  
  mysis_length_percentile_outlier_df  <-rbind(mysis_length_percentile_outlier_df,
                                              OneRowofOutlier)
  }

```

## Generalized ESD Test for Outliers

```{r Rosner's GESD Test for length Outliers}
mysis_length_rosner_outlier_df <- data.frame()
gemma_length_rosner_outlier_df <- data.frame()

#iterate through age categories and create temp df for that age category
for (AgeCategory in levels(growth_gemma_outlier_df$Age_Category)){
  AgeGroup_df <- growth_gemma_outlier_df %>%
  filter(Age_Category == AgeCategory)
  
  gemma_rosner_length <- rosnerTest(AgeGroup_df$Length_mm, k=10, alpha = 0.05)

  Rosner_Outliers <- gemma_rosner_length[["all.stats"]] %>%
    filter(Outlier == TRUE)
  
  gemma_rosner_outlier_rownum <- AgeGroup_df %>% 
                                filter(row_number() %in% Rosner_Outliers$Obs.Num)

  gemma_length_rosner_outlier_df <-rbind(gemma_length_rosner_outlier_df,
                                         gemma_rosner_outlier_rownum)
}

for (AgeCategory in levels(growth_mysis_outlier_df$Age_Category)){
  AgeGroup_df <- growth_mysis_outlier_df %>%
  filter(Age_Category == AgeCategory)

  mysis_rosner_length <- rosnerTest(AgeGroup_df$Length_mm, k=10, alpha = 0.05)

  Rosner_Outliers <- mysis_rosner_length[["all.stats"]] %>%
    filter(Outlier == TRUE)

  mysis_rosner_outlier_rownum <- AgeGroup_df %>% 
                                    filter(row_number() %in% Rosner_Outliers$Obs.Num)

  mysis_length_rosner_outlier_df <-rbind(mysis_length_rosner_outlier_df, 
                                         mysis_rosner_outlier_rownum)
}

```


## Cooks Distance Outliers

```{r Cooks Distance Outliers}
mysis_length_cooksd_outlier_df <- data.frame()
gemma_length_cooksd_outlier_df <- data.frame()

for (AgeCategory in levels(growth_mysis_outlier_df$Age_Category)){
  AgeGroup_df <- growth_mysis_outlier_df %>%
  filter(Age_Category == AgeCategory)

  model <- lm(Mass_mg ~ Length_mm, data = AgeGroup_df)
  cooksD <- cooks.distance(model)
  influence_cuttoff <- 3 * mean(cooksD, na.rm =TRUE)
  influencial_Observations <- cooksD[(cooksD > influence_cuttoff)]
 
  plot(cooksD, pch="*", cex=2, main = "Influential Obs by Cooks distance") 
  abline(h = influence_cuttoff,col="red")
  text(x=1:length(cooksD) + 1, 
       y=cooksD, 
       labels = ifelse(cooksD > influence_cuttoff, names(cooksD),""),
       col="red")
 
  mysis_cooksd_outlier_rownum <- AgeGroup_df %>% filter(row_number() %in% 
                                                       names(influencial_Observations))

  mysis_length_cooksd_outlier_df <-rbind(mysis_length_cooksd_outlier_df,
                                         mysis_cooksd_outlier_rownum)
}

for (AgeCategory in levels(growth_gemma_outlier_df$Age_Category)){
  AgeGroup_df <- growth_gemma_outlier_df %>%
  filter(Age_Category == AgeCategory)

  model <- lm(Mass_mg ~ Length_mm, data = AgeGroup_df)
  cooksD <- cooks.distance(model)
  influence_cuttoff <- 3 * mean(cooksD, na.rm =TRUE)
  influencial_Observations <- cooksD[(cooksD > influence_cuttoff)]

  plot(cooksD, pch="*", cex=2, main = "Influential Obs by Cooks distance") 
  abline(h = influence_cuttoff,col="red")
  text(x=1:length(cooksD) + 1, 
       y=cooksD, 
       labels = ifelse(cooksD > influence_cuttoff, names(cooksD),""),
       col="red")

  gemma_cooksd_outlier_rownum <- AgeGroup_df %>% filter(row_number() %in% 
                                                          names(influencial_Observations))

  gemma_length_cooksd_outlier_df <-rbind(gemma_length_cooksd_outlier_df,
                                         gemma_cooksd_outlier_rownum)
}

```



## Length Outlier Comparison

```{r length Outlier Summary}

#create summary dataframe of all possible outliers using a list of dataframes for antijoin
mysis_length_summary_outlier_df <- mysis_length_boxplot_outlier_df
gemma_length_summary_outlier_df <- gemma_length_boxplot_outlier_df

mysis_list_outlier_df <- list(mysis_length_boxplot_outlier_df, mysis_length_percentile_outlier_df,
                              mysis_length_rosner_outlier_df, mysis_length_cooksd_outlier_df)
gemma_list_outlier_df <- list(gemma_length_boxplot_outlier_df, gemma_length_percentile_outlier_df,
                              gemma_length_rosner_outlier_df, gemma_length_cooksd_outlier_df)  

for (outlier_dfs in mysis_list_outlier_df) {
  Outliers_notin_summary_df <- anti_join(outlier_dfs, mysis_length_summary_outlier_df)

  mysis_length_summary_outlier_df <- rbind(mysis_length_summary_outlier_df, 
                                          Outliers_notin_summary_df)
}

for (outlier_dfs in gemma_list_outlier_df) {
  Outliers_notin_summary_df <- anti_join(outlier_dfs, gemma_length_summary_outlier_df)
 
  gemma_length_summary_outlier_df <- rbind(gemma_length_summary_outlier_df,
                                          Outliers_notin_summary_df)
}


#Wrangle Dataframes
mysis_length_summary_outlier_df <- mysis_length_summary_outlier_df %>% arrange(Sample_ID)
gemma_length_summary_outlier_df <- gemma_length_summary_outlier_df %>% arrange(Sample_ID)

row.names(mysis_length_summary_outlier_df) <- NULL
row.names(gemma_length_summary_outlier_df) <- NULL


mysis_length_outlier_method_df <- subset(mysis_length_summary_outlier_df,
                                        select=-c(Date, Tank_Location, 
                                                  Feed_Group, Mass_mg, Mass_Outlier)) 
gemma_length_outlier_method_df <- subset(gemma_length_summary_outlier_df,
                                        select=-c(Date, Tank_Location, 
                                                  Feed_Group, Mass_mg, Mass_Outlier)) 

#add columns for each method
methods_list = list("Boxplot", "Percentile", "Rosner", "CooksD")

for (method in methods_list){
  mysis_length_outlier_method_df <- mysis_length_outlier_method_df %>% 
    add_column(!!(method) :=0)#add column for each list and set value to 0
  gemma_length_outlier_method_df <- gemma_length_outlier_method_df %>%
    add_column(!!(method) :=0)
}

# for each outlier, add +1 to any method column that detected that outlier. 
###Mysis###
for (sample_ID in mysis_length_outlier_method_df$Sample_ID){
  if (sample_ID %in% mysis_length_boxplot_outlier_df$Sample_ID){
    mysis_length_outlier_method_df["Boxplot"][mysis_length_outlier_method_df["Sample_ID"] ==
                                                sample_ID] <- 1
  }
  if (sample_ID %in% mysis_length_percentile_outlier_df$Sample_ID){
    mysis_length_outlier_method_df["Percentile"][mysis_length_outlier_method_df["Sample_ID"] == 
                                                   sample_ID] <-1
  }
  if (sample_ID %in% mysis_length_rosner_outlier_df$Sample_ID){
    mysis_length_outlier_method_df["Rosner"][mysis_length_outlier_method_df["Sample_ID"] == 
                                               sample_ID] <-1
  }
  if (sample_ID %in% mysis_length_cooksd_outlier_df$Sample_ID){
    mysis_length_outlier_method_df["CooksD"][mysis_length_outlier_method_df["Sample_ID"] == 
                                               sample_ID] <-1
  }
}

###Gemma###
for (sample_ID in gemma_length_outlier_method_df$Sample_ID){
  if (sample_ID %in% gemma_length_boxplot_outlier_df$Sample_ID){
    gemma_length_outlier_method_df["Boxplot"][gemma_length_outlier_method_df["Sample_ID"] == 
                                                sample_ID] <-1
  }
  if (sample_ID %in% gemma_length_percentile_outlier_df$Sample_ID){
    gemma_length_outlier_method_df["Percentile"][gemma_length_outlier_method_df["Sample_ID"] == 
                                                   sample_ID] <-1
  }
  if (sample_ID %in% gemma_length_rosner_outlier_df$Sample_ID){
    gemma_length_outlier_method_df["Rosner"][gemma_length_outlier_method_df["Sample_ID"] == 
                                               sample_ID] <-1
  }
  if (sample_ID %in% gemma_length_cooksd_outlier_df$Sample_ID){
    gemma_length_outlier_method_df["CooksD"][gemma_length_outlier_method_df["Sample_ID"] == 
                                               sample_ID] <-1
  }
}

#Create a total column of outlier methods, to sum all detections among methods per sample

mysis_length_outlier_method_df <- mysis_length_outlier_method_df %>%
 mutate(Total = select(., Boxplot:CooksD) %>% rowSums(na.rm =TRUE))

gemma_length_outlier_method_df <- gemma_length_outlier_method_df %>%
 mutate(Total = select(., Boxplot:CooksD) %>% rowSums(na.rm =TRUE))

#Add a length_outlier column to Mysis Outlier Dataframe. Transfer Total Method data 
#to Length outlier samples. Insert 0, where no outliers detected.

for (Sample in growth_mysis_outlier_df$Sample_ID) {
  if (Sample %in% mysis_length_outlier_method_df$Sample_ID){
    growth_mysis_outlier_df$length_Outlier[growth_mysis_outlier_df["Sample_ID"] == Sample] <- 
      mysis_length_outlier_method_df$Total[mysis_length_outlier_method_df["Sample_ID"] == Sample]
  } 
  else{
    growth_mysis_outlier_df$length_Outlier[growth_mysis_outlier_df["Sample_ID"] == Sample] <- 0
  }
}

for (Sample in growth_gemma_outlier_df$Sample_ID) {
  if (Sample %in% gemma_length_outlier_method_df$Sample_ID){
    growth_gemma_outlier_df$length_Outlier[growth_gemma_outlier_df["Sample_ID"] == Sample] <- 
      gemma_length_outlier_method_df$Total[gemma_length_outlier_method_df["Sample_ID"] == Sample]
  } 
  else{
    growth_gemma_outlier_df$length_Outlier[growth_gemma_outlier_df["Sample_ID"] == Sample] <- 0
  }
}

growth_mysis_outlier_df$length_Outlier <- factor(growth_mysis_outlier_df$length_Outlier, 
                                 levels = c("4","3","2","1","0"))

growth_gemma_outlier_df$length_Outlier <- factor(growth_gemma_outlier_df$length_Outlier, 
                                 levels = c("4","3","2","1","0"))

#plot outlier frequency based on method selection
Mysis_LMR_Scatter_Length_Outliers <- ggplot(growth_mysis_outlier_df, 
       aes(x = Length_mm, y = Mass_mg, color = length_Outlier)) +
  geom_point() + 
  scale_color_brewer(palette="Set2") 

Gemma_LMR_Scatter_Length_Outliers <- ggplot(growth_gemma_outlier_df, 
       aes(x = Length_mm, y = Mass_mg, color = length_Outlier)) +
  geom_point() +
  scale_color_brewer(palette="Set2") 

Mysis_Mass_Boxplot_Length_Outliers <- ggplot(growth_mysis_outlier_df, 
       aes(y=Length_mm, x=Age_Category)) + 
  geom_boxplot() + 
  geom_point(aes(color = length_Outlier)) +
  scale_color_brewer(palette="Set1") + 
  theme_classic()

Gemma_Mass_Boxplot_Length_Outliers <- ggplot(growth_gemma_outlier_df, 
       aes(y=Length_mm, x=Age_Category)) + 
  geom_boxplot() + 
  geom_point(aes(color = length_Outlier)) +
  scale_color_brewer(palette="Set1") + 
  theme_classic()
```

# Summarize Outlier Testing
Now that I have an understanding of different methods for outlier collection,
I was curious how  these methods would compare between variables. 
Which outliers would recieve marks from the most methods for detection? 
Which would only recieve one mark? 
Do the Mass outliers correspond to the Length Outliers?

```{r Difference Between Outliers By Length or Mass}

#Tidy the data
growth_mysis_outlier_tidy_df<-growth_mysis_outlier_df %>%
  pivot_longer(c("Mass_Outlier","length_Outlier"), names_to = "Outliers_Metric", values_to = "Outlier_Count")

#difference in # of samples from both length & mass that each had 1,2,3,4 method counts 
growth_mysis_outlier_tidy_df %>%
ggplot(., aes(x=Outlier_Count,fill=Outliers_Metric ))+geom_bar()

#counts between length and mass of samples that had 0,1,2,3,4 methods 
ggplot(growth_mysis_outlier_tidy_df, aes(x=Outliers_Metric, fill=Outlier_Count ))+geom_bar()


#Counts for total outliers of length and total outliers for mass
total_length_Outlier_Count <- growth_mysis_outlier_tidy_df %>%
  filter(., Outlier_Count != 0)%>%
  filter(.,Outliers_Metric == "length_Outlier") %>%
  nrow()

total_mass_Outlier_Count <- growth_mysis_outlier_tidy_df %>%
  filter(., Outlier_Count != 0)%>%
  filter(.,Outliers_Metric == "Mass_Outlier") %>%
  nrow()


df_mysis_mass_one_method<-mysis_mass_outlier_method_df %>%
  filter(Sample_ID %in% df_outliers_one_method$Sample_ID)
df_mysis_length_one_method<-mysis_length_outlier_method_df %>%
  filter(Sample_ID %in% df_outliers_one_method$Sample_ID)

#Using count we will normalize the bar plot so it shows % of # of methods
growth_mysis_outlier_tidy_df %>%
ggplot(., aes(x=Outlier_Count,fill=Outliers_Metric ))+geom_bar()

#scatter plot of length v mass with mass outliers 
growth_mysis_outlier_df %>%
ggplot(., aes(y=Mass_mg, x=Length_mm)) + 
  geom_point(aes(color=Mass_Outlier))
#scatter plot of length v mass with mass outliers 
growth_mysis_outlier_df %>%
ggplot(., aes(y=Mass_mg, x=Length_mm)) + 
  geom_point(aes(color= length_Outlier))

#plotting outliers that received one method of detection for mass
ggplot(growth_mysis_nofactor_outlier_df, aes(y=Mass_mg,x=Age_Category)) + 
         geom_boxplot()+scale_color_brewer(palette="Set1") + 
  geom_jitter(data=. %>% filter(Sample_ID %in% df_mysis_mass_one_method$Sample_ID),color="red")                                                                 
#plotting outliers that received one method of detection for length

ggplot(growth_mysis_nofactor_outlier_df, aes(y=Length_mm,x=Age_Category)) + 
         geom_boxplot()+scale_color_brewer(palette="Set1") + 
  geom_point(data=. %>% filter(Sample_ID %in% df_mysis_length_one_method$Sample_ID),color="red")                                                                 

```

```{r Combined Length and Mass Outliers}

growth_mysis_outlier_length_mass_df<- growth_mysis_outlier_df
growth_gemma_outlier_length_mass_df <- growth_gemma_outlier_df

#factors have to be converted to strings first, or the factor level will transform as numeric
growth_mysis_outlier_length_mass_df$Mass_Outlier <- as.numeric(as.character(growth_mysis_outlier_length_mass_df$Mass_Outlier))
growth_mysis_outlier_length_mass_df$length_Outlier <- as.numeric(as.character(growth_mysis_outlier_length_mass_df$length_Outlier))
#gemma
growth_gemma_outlier_length_mass_df$Mass_Outlier <- as.numeric(as.character(growth_gemma_outlier_length_mass_df$Mass_Outlier))
growth_gemma_outlier_length_mass_df$length_Outlier <- as.numeric(as.character(growth_gemma_outlier_length_mass_df$length_Outlier))

#create a sum of both length and mass outliers to see which observations are most impacted
growth_mysis_outlier_length_mass_df <- growth_mysis_outlier_length_mass_df %>%
 mutate(Total_Outlier = select(., Mass_Outlier:length_Outlier) %>% rowSums(na.rm =TRUE))

growth_gemma_outlier_length_mass_df <- growth_gemma_outlier_length_mass_df %>%
 mutate(Total_Outlier = select(., Mass_Outlier:length_Outlier) %>% rowSums(na.rm =TRUE))

#refactor the total to get discrete values for ploting

growth_mysis_outlier_length_mass_df$Total_Outlier <- factor(growth_mysis_outlier_length_mass_df$Total_Outlier, 
                                 levels = c("8","7","6","5","4","3","2","1","0"))

growth_gemma_outlier_length_mass_df$Total_Outlier <- factor(growth_gemma_outlier_length_mass_df$Total_Outlier, 
                                 levels = c("7","6","5","4","3","2","1","0"))

#which method gives us the 1's in total outliers for both length and mass combined
df_mysis_outliers_one_method <- growth_mysis_outlier_length_mass_df %>%
  filter(Total_Outlier==1)

#Plotting only te outlines that received 1 method for both length and mass combined
ggplot(growth_mysis_nofactor_outlier_df, aes(x=Length_mm,y=Mass_mg, color=Total_Outlier==1)) +geom_point() + 
  
  scale_color_brewer(palette="Dark2") +
  labs(title='Length by Age, Tank Location, & Feed Type', 
       subtitle = "Box Plot",
       y = "Length (mm)", x = "Age Category") +
  theme_minimal() +
  theme(
    axis.title = element_text(size = 30),
    axis.text = element_text(size = 20),
    plot.title = element_text(size = 40, face = "bold"),
    plot.subtitle = element_text(size = 25),
    legend.title = element_text(size =20),
    legend.text = element_text(size =15),
    legend.key.size = unit(2,"cm"),
    strip.text.x = element_text(size=20)
    )

#Using the summary of values that were flagged as outliers in both length and mass features, I 
#create a scatter plot that indicates where these outliers are located and how many of the 
#methods detected said value (e.g. 8 means the values was flagged by 4 features in mass and 4 features for length)
M_O_S <- ggplot(growth_mysis_outlier_length_mass_df, aes(y=Mass_mg, x=Length_mm, color= Total_Outlier ))+ 
  geom_point(size=2)+
  scale_color_brewer(palette="Set1")+
  ylim(y=c(0,7000))+xlim(c(20,65))+
  labs(title='Mysis Length vs. Mass Highlighting Outliers',
       subtitle='Outlier detection count for both mass and length',
       y = "Mass (mg)", x = "Length (mm)") +
  theme_minimal() +
  theme(
    axis.title = element_text(size = 30),
    axis.text = element_text(size = 20),
    plot.title = element_text(size = 40, face = "bold"),
    plot.subtitle = element_text(size = 25),
    legend.title = element_text(size =25),
    legend.text = element_text(size =20),
    legend.key.size = unit(2,"cm"),
    strip.text.x = element_text(size=20)
    )


#Using the summary of values that were flagged as outliers in both length and mass features, I 
#create a scatter plot that indicates where these outliers are located and how many of the 
#methods detected said value (e.g. 8 means the values was flagged by 4 features in mass and 4 features for length)
G_O_S <- ggplot(growth_gemma_outlier_length_mass_df, aes(y=Mass_mg, x=Length_mm, color= Total_Outlier ))+
   geom_point(size=2)+
  scale_color_brewer(palette="Dark2")+
  ylim(y=c(0,7000))+xlim(c(20,65))+
  labs(title='Gemma Length vs. Mass Highlighting Outliers',
       subtitle='Outlier detection count for both mass and length',
       y = "Mass (mg)", x = "Length (mm)") +
  theme_minimal() +
  theme(
    axis.title = element_text(size = 30),
    axis.text = element_text(size = 20),
    plot.title = element_text(size = 40, face = "bold"),
    plot.subtitle = element_text(size = 25),
    legend.title = element_text(size =25),
    legend.text = element_text(size =20),
    legend.key.size = unit(2,"cm"),
    strip.text.x = element_text(size=20)
    )
```


```{r Medium Visuals}


#saving png files for medium post
pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/Mysis_Outlier_summary.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(M_O_S)
invisible(dev.off())
knitr::include_graphics(pngfile)

pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/Gemma_Outlier_summary.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(G_O_S)
invisible(dev.off())
knitr::include_graphics(pngfile)


```
