---
title: "4.0 Modeling"
author: "Andrew Ingalls"
date: "3/29/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)
library(viridis) #coloring for ridge plots
library(ggResidpanel) # panel of diagnostic plots including cooksd
library(qqplotr) #qq plots
library(rstatix) #t-test, wilcoxon, anova, kruskal-wallis, correlation, levenetest,
library(broom) #output tibbles for statistical modeling 
library(ggpubr) #publication ready plots
library(ragg) #help with png creation
library(gridExtra) #allows for grid.arrange functionality
library(WRS2) #collection of robust statistical methods based on Wilcox;
library(emmeans) #pairwise

```

# Growth

## Mass Significance Testing by Age for Feed Groups
Because we don't have strictly normal distributions for our data, I won't be using
any tests that assume normality. To confirm our intuitions about a lack of significant
difference in growth between the Feed Groups, we will have to test the 
null hypothesis (p-value < 0.05). Our null hypothesis is: There is no difference in Mass
between the Feed Group for each Age group. 

First we will take a look at the variances between the populations. This will
help us further understand what type of significance test to perform.

Since we aren't confident in normality of our test, we will use the Levene's test
to test the variance. 

For Significance, we are curious if either of the Feed Groups has a superior performance,
so we will use a two-tailed test.

```{r Boxplot Significance Testing}
#Variance for Mass for Age Groups between Feed Groups are not Significantly Different
df_growth %>%
  group_by(Age_Category)%>%
  levene_test(Mass_mg ~ Feed_Group)

#Two Sample unpaired nonparametric test. iterate over all age categories
Wilcox_Mass_Age_P_Table <- data.frame(matrix(nrow = 0, ncol = 2))
colnames(Wilcox_Mass_Age_P_Table) <- c('Age Category','P-Value')

for (AgeCategory in levels(df_growth$Age_Category)){
  Mysis_Age_df_filter <- df_growth%>% 
                            filter(., Age_Category == AgeCategory & 
                                      df_growth$Feed_Group == 'Mysis')
  Gemma_Age_df_filter <- df_growth %>% 
                            filter(., Age_Category == AgeCategory &
                                      df_growth$Feed_Group == 'Gemma')
  
  SigTest <- wilcox.test(Mysis_Age_df_filter$Mass_mg, Gemma_Age_df_filter$Mass_mg, 
                         alternative = "two.sided", paired = FALSE)
  
  #add new significance data one row below previous
  Wilcox_Mass_Age_P_Table[nrow(Wilcox_Mass_Age_P_Table) + 1, ] <- c(AgeCategory, 
                                                                   SigTest$p.value)
}
#No significant difference between the Feed Groups for any Age Category
Mass_boxplot <- ggboxplot(df_growth, x="Age_Category", y="Mass_mg",
                          color = "Feed_Group", add = "jitter", 
                          lwd = 2, fill = 'grey', outlier.colour = NULL, 
                          outlier.size = 3) 

Mass_P_Graph <- Mass_boxplot + 
  stat_compare_means(aes(group = Feed_Group), label = "p.format", size = 7) + 
  scale_color_brewer(palette = "Dark2") +
  labs(title = 'Mass of A. mexicanus by Age & Feed Type', 
       subtitle = "Wilcoxon Rank Sum Test",
       y = "Mass (mg)", x = "Age Category") +
  theme_specs

```

## Length Significance Testing by Age for Feed Groups

```{r}
#Variance for Length for Age Groups between Feed Groups 
#are significantly different @ three months
df_growth %>%
  group_by(Age_Category)%>%
  levene_test(Length_mm ~ Feed_Group)

#Two Sample unpaired nonparametric test. iterate over all age categories
Wilcox_Length_Age_P_Table <- data.frame(matrix(nrow = 0, ncol = 2))
colnames(Wilcox_Length_Age_P_Table) <- c('Age Category', 'P-Value')


for (AgeCategory in levels(df_growth$Age_Category)){
  Mysis_Age_df_filter <- df_growth%>% 
                            filter(., Age_Category == AgeCategory & 
                                      df_growth$Feed_Group == 'Mysis')
  Gemma_Age_df_filter <- df_growth %>% 
                            filter(., Age_Category == AgeCategory &
                                      df_growth$Feed_Group == 'Gemma')
  
  SigTest <- wilcox.test(Mysis_Age_df_filter$Length_mm, Gemma_Age_df_filter$Length_mm, 
                         alternative = "two.sided", 
                         paired= FALSE)
  
  Wilcox_Length_Age_P_Table[nrow(Wilcox_Length_Age_P_Table) + 1, ] <- c(AgeCategory, 
                                                                       SigTest$p.value)
}#No significant difference between the Feed Groups for any Age Category

Length_box_plot <- ggboxplot(df_growth, x="Age_Category", y="Length_mm",
                             color = "Feed_Group", add = "jitter",
                             lwd = 2, fill = 'grey', outlier.colour = NULL, 
                             outlier.size = 3) 

Length_P_Graph <- Length_box_plot + 
  stat_compare_means(aes(group = Feed_Group), label = "p.format", size = 7) + 
  scale_color_brewer(palette = "Dark2") +
  labs(title = 'Length of A. mexicanus by Age & Feed Type', 
       subtitle = "Wilcoxon Rank Sum Test",
       y = "Length (mm)", x = "Age Category") +
  theme_specs
```

## Length-Mass Relationship Modeling for Feed Groups 
Based on our distribution plots, we can be pretty certain a linear model is not appropriate
for our models. However, we will use the Residual plot to help us understand our model better.
Using these plots its pretty clear a polynomial model is more appropriate. 

This is important in the husbandry of fish as this relationship gives us the growth of the fish,
its health, and a realative condition within its life.

```{r Residual Plots}
y <- df_growth$Mass_mg
x <- df_growth$Length_mm

mod <- lm(y~x)

Growth_resid <- resid_panel(mod, plots='resid') +theme_specs

```

```{r Transform Data for Linear Regression}

df_growth$Mass_log <- log(df_growth$Mass_mg)
df_growth$Length_log <- log(df_growth$Length_mm)
y2 <- df_growth$Mass_log
x2 <- df_growth$Length_log

mod_log <- lm(y2~x2)


resid_panel(mod_log,plot='resid')

shapiro.test(resid(mod_log))

#Compare the linear model to a log transformed model 
Compare_log_resid <- resid_compare(models=list(mod,
                          mod_log),
              plots= c("resid", "qq"),
              smoother = TRUE,
              qqbands = TRUE,
              title.opt = FALSE)
#Log transform data looks like a great fit


by_feedgroups <- group_by( df_growth, Feed_Group)

do(by_feedgroups,tidy(lm(Mass_log~Length_log,data=.)))


model_FeedGroups <- lm(Mass_log~Length_log*Feed_Group, df_growth)

summary(model_FeedGroups)

glance(model_FeedGroups)


plot(df_growth$Length_log, df_growth$Mass_log, col='red', main='Summary of Regression Model', xlab='x', ylab='y')


#The model is statistically significant, however the groups are not. Very tight R2

summary(aov(Mass_log ~ Length_log*Feed_Group, data=df_growth))

```

```{r Correlation and Plotting of Each Feed Group}

#Spearman Correlation, which does not require normality
df_growth_mysis <- df_growth %>%
  filter(.,Feed_Group=="Mysis")  
cor(df_growth_mysis$Length_mm,df_growth_mysis$Mass_mg, method="spearman")

df_growth_gemma <- df_growth %>%
  filter(.,Feed_Group=="Gemma")  
cor(df_growth_gemma$Length_mm,df_growth_gemma$Mass_mg, method="spearman")

#Visual representation of spearman correlation  
Mysis_Spearman <- ggscatter(df_growth_mysis, x='Length_log', y='Mass_log', add = "reg.line", conf.int=TRUE, 
          cor.coef = TRUE, cor.method = "spearman", xlab = "Length (mm)", ylab = "Mass (mg)",
          title = "Mysis: Length Mass Log Relationship", subtitle = "Spearman Correlation")

Gemma_Spearman <- ggscatter(df_growth_gemma, x='Length_log', y='Mass_log', add = "reg.line", conf.int=TRUE, 
          cor.coef = TRUE, cor.method = "spearman", xlab = "Length (mm)", ylab = "Mass (mg)",
          title = "Gemma: Length Mass Log Relationship", subtitle = "Spearman Correlation")


#Create linear model for each feed group using log transformations.
mysis_linear_model <- lm(Mass_log~Length_log,data=df_growth_mysis)
gemma_linear_model <- lm(Mass_log~Length_log,data=df_growth_gemma) 

#Residuals % Q-Q plots
plot(mysis_linear_model)
plot(gemma_linear_model)

```

```{r Plot Feed Groups with poly and linear line for visualization}

#Linear Relationship Assumption
Linear_relation <- ggscatter(
  df_growth, x = 'Length_log' , y =  'Mass_log',
  color = "Feed_Group", add = "reg.line"
  )+
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~"), color = Feed_Group), size = 10
    )  +
  scale_color_brewer(palette="Dark2") +
  scale_fill_brewer(palette="Dark2") +
  labs(title='Length-Mass Relationship of A. mexicanus by Feed Type', 
       subtitle = "Log-Log Linear Regression Model", 
       y = "log(Mass (mg))", x = "log(Length (mm))") +
  theme_specs

#Homogeneity of regression slopes 
#Feed_Group interaction was not statisticaly significant F(1,730) = 0.390 p = 0.533
df_growth %>% anova_test(Mass_log ~ Feed_Group*Length_log)

#Normality of Residuals
feed_model <-  lm(Mass_log ~ Length_log + Feed_Group, data =df_growth)
feed_model.metrics <- augment(feed_model) %>%
  select(-.hat,-.sigma,-.fitted) #remove unneeded details

head(feed_model.metrics,3)

shapiro_test(feed_model.metrics$.resid)

#shapiro test indicated the model is not from a normal distribution


#Wilcox ancova to compare two linear regression lines
#only significantly different on the extreme ends of the regression model

ancova(Mass_mg ~ Length_mm + Feed_Group, df_growth)


```


# Survival

## Signficance Testing between Census Events for Census and Deaths

```{r  Survival}
Wilcox_census <- data.frame(matrix(nrow=0, ncol = 2))
colnames(Wilcox_census) <-c('Sample Date','P_Value')

for (Date in levels(factor(df_survival$Date))){
   if (Date == "2020-10-15"){
     next
     }else{
  Mysis_Census <- df_survival$Census[df_survival$Date == Date & 
                         df_survival$Feed_Group == 'Mysis']

  Gemma_Census <- df_survival$Census[df_survival$Date == Date &
                         df_survival$Feed_Group == 'Gemma']
  
  SigTest <- wilcox.test(Mysis_Census, Gemma_Census)
  
  Wilcox_census[nrow(Wilcox_census) + 1,] <- c(Date, SigTest$p.value)
     }
}

Census_Wilcox <- ggline(df_survival, x = "Date", y = "Census", add = "mean_se",
          color = "Feed_Group") +
  stat_compare_means(aes(group = Feed_Group), label = "p.format", size = 7) +
  scale_color_brewer(palette="Dark2") +
  scale_fill_brewer(palette="Dark2") +
  labs(title='Census over time of A. mexicanus by Feed Type', 
       subtitle = "Wilcoxon Rank Sum Test", 
       y = "Census", x = "Sampling Date") +
  theme_specs



#Deaths

df_deaths <- df_survival %>%
  group_by(Tank_Location, Feed_Group) %>%
  summarize(Census_Dead = -diff(Census),
            Date = Date[-1])

Wilcox_deaths <- data.frame(matrix(nrow=0, ncol = 2))
colnames(Wilcox_deaths) <-c('Sample Date','P_Value')

for (Date in levels(factor(df_deaths$Date))){

  Mysis_death <- df_deaths$Census_Dead[df_deaths$Date == Date & 
                         df_deaths$Feed_Group == 'Mysis']

  Gemma_death <- df_deaths$Census_Dead[df_deaths$Date == Date &
                         df_deaths$Feed_Group == 'Gemma']
  
  SigTest <- t.test(Mysis_death, Gemma_death)
  
  Wilcox_deaths[nrow(Wilcox_deaths) + 1,] <- c(Date, SigTest$p.value)
     }

df_deaths$Date= factor(df_deaths$Date)

wilcox_deaths <- ggbarplot(df_deaths, x = "Date", y = "Census_Dead", add = "mean_se",
          color = "Feed_Group", fill = "Feed_Group", position = position_dodge(0.8))+
  stat_compare_means(aes(group = Feed_Group), label = "p.format", size = 7)+
  scale_color_brewer(palette="Dark2") +
  scale_fill_brewer(palette="Dark2") +
  labs(title='Deaths between Events of A. mexicanus by Feed Type', 
       subtitle = "Wilcoxon Rank Sum Test", 
       y = "Deaths", x = "Sampling Date") +
  theme_specs


Wilcox_census
Wilcox_deaths
```


```{r  Survival Model}
#linear model had very low R-squared value (<0.45 & <0.2), switched to polynomial
df_survival$Date= as.Date(df_survival$Date)

formula <- y~poly(x, 2, raw = TRUE)

p<- ggplot(df_survival, aes(Date, Census, color = Feed_Group)) + 
  geom_point() +
  stat_smooth(aes(fill = Feed_Group, color=Feed_Group), method='lm', formula = formula)+
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~")),
    formula = formula ) +  
  theme_bw()
ggpar(p)

#poor fit

```

# Sex Sort

##Significance Testing between Feed Groups


```{r Sex Sort}

#Create a column of total counts for each tank location
df_sex_sort$Total <- rowSums(df_sex_sort[,c("Female","Male","Unknown")])

df_sex_norm <- df_sex_sort %>%
  group_by(Feed_Group)%>%
  summarise(Female=Female/Total,Male=Male/Total,Unknown=Unknown/Total)

df_sex_tidy <- df_sex_norm %>% pivot_longer(Female:Unknown, names_to = "Sex", values_to = "Percentage")

#Still using Wilcoxon Rank Sum, no significant difference in Sex between groups
Sex_wilcox <- ggboxplot(df_sex_tidy, x= "Sex", y="Percentage", color = "Feed_Group", fill = 'grey') +
  stat_compare_means(aes(group = Feed_Group),label = "p.format", size = 7) +
  scale_color_brewer(palette="Dark2") +
  scale_fill_brewer(palette="Dark2") +
  labs(title='Sex Sorting of A. mexicanus by Feed Type', 
       subtitle = "Wilcoxon Rank Sum Test", 
       y = "Sex Count", x = "Sex Type") +
  theme_specs



```

# Fecundity

## Significance Testing between Feed Groups

```{r Fecundity}
wilcox.test(Viable_Embryos ~ Feed_Group, data=df_fecundity)

Fecundity_Wilcox <- ggboxplot(df_fecundity, x= "Feed_Group", y="Viable_Embryos", color = "Feed_Group", fill = "grey") +
  stat_compare_means(size = 7)+
  scale_color_brewer(palette="Dark2") +
  scale_fill_brewer(palette="Dark2") +
  labs(title='Sex Sorting of A. mexicanus by Feed Type', 
       subtitle = "Wilcoxon Rank Sum Test", 
       y = "Sex Count", x = "Sex Type") +
  theme_specs



```

#Medium Plots
```{r Medium Plots}

#Growth

pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/Sig_Mass.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Mass_P_Graph)
invisible(dev.off())
knitr::include_graphics(pngfile)


pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/Sig_Length.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Length_P_Graph)
invisible(dev.off())
knitr::include_graphics(pngfile)

pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/Resid_growth.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Growth_resid)
invisible(dev.off())
knitr::include_graphics(pngfile)

pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/Resid_compare.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Compare_log_resid)
invisible(dev.off())
knitr::include_graphics(pngfile)

pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/Mysis_spearman.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Mysis_Spearman)
invisible(dev.off())
knitr::include_graphics(pngfile)

pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/Gemma_Spearman.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Gemma_Spearman)
invisible(dev.off())
knitr::include_graphics(pngfile)

pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/Growth_Linear.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Linear_relation)
invisible(dev.off())
knitr::include_graphics(pngfile)

##Survival##
pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/Census_Wilcox.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Census_Wilcox)
invisible(dev.off())
knitr::include_graphics(pngfile)

pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/Deaths_Wilcox.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(wilcox_deaths)
invisible(dev.off())
knitr::include_graphics(pngfile)


##Sex##
pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/Sex_Wilcox.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Sex_wilcox)
invisible(dev.off())
knitr::include_graphics(pngfile)

##Fecundity##
pngfile <- fs::path(knitr::fig_path(),  "analysis/EDA_Figures/fECUNDITY_Wilcox.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300)
plot(Fecundity_Wilcox)
invisible(dev.off())
knitr::include_graphics(pngfile)

```

